\chapter{Section 5
Practical Example of self care}

%Relief- Titled: Letting it go

I almost always feel physical and emotional relief drawing a mandala. It is astounding how much I feel I have healed with each line of the image and as well as with the completion of each mandala. As I have mentioned previously, I have kept a mandala diary over the last 2 years and this is a small example of the self care I have undertaken. The next four mandalas depict the exploration of betrayal and dishonesty. As a result, I was able to let go of the emotional wounds I had been suffering. I feel that this process has liberated me from the destruction that I once felt in relationship to my former partner. 


%FIGURE 2-KEPT IN THE DARK


I drew this mandala when I found out from a friend that my former partner married one of our Khmer friends who was his assistant at work. It was 1 year and 7 months after our relationship ended. 


%FIGURE 3-PISSED OFF

Initially, I was very angry with the deceit and lies of my former partner. 

%FIGURE 4-HURT OTHERS COULD NOT BE HONEST

I felt hurt that there where people around me who knew the situation but said nothing. 

%FIGURE 5- LETTING GO OF THE BULLSHIT AND REALIZING THERE IS SO MUCH MORE OUT THERE


By the 4th mandala, I felt a huge relief which was a transition from being so angry at the beginning of drawing these 4 mandalas to letting go and moving on. The symbolic nature of this 4th mandala prompted me to think of the universe and appreciate the wider view of my circumstances. It is fascinating to note that I was not affected by my former partner marriage, but only dismayed about the lies and disrespect he showed towards me. 

At the beginning of my thesis all I wanted was clarity and to understand why he chose to leave, but amazingly through this inquiry I was able to create my own clarity in terms of my life without him, which feels much more powerful and empowering than I could have ever hoped for. 



